# Correct data {#correct-data}
First, we define functions to 1) get background-corrected values of our samples, 2) compute the standard concentrations, and 3) to get tidy data frames of the standard and the biological samples.
```{r, corr-data, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE}
# Gets samples, performs background correction
# Works only on single rows of the 96-well plate
# The number of columns is user defined (set n.cols)
#' @param data input data frame
#' @param row.idx row index in the input data frame of the sample
#' @param col.dix column index in the input data frame of the first cell of the sample
#' @param n.cols number of columns to get in the input data frame
#' @return a data frame with background-corrected values
correct.sample <- function(data, row.idx, col.idx, n.cols) {
  row.end <- row.idx + 1 # we select only two rows (450 nm and 620 nm measurements)
  col.end <- col.idx + (n.cols - 1) # the number of columns is user defined
  sample <- data %>%
    .[row.idx:row.end, col.idx:col.end]
  corr.sample <- sample[1, ] - sample[2, ]
  return(as.tibble(corr.sample))
}

# Compute dilution series
#' @param max.conc.std maximal concentration
#' @param n.dilutions number of dilutions performed
#' @param dilution.fact dilution factor
#' @return a vector with concentrations
comp.dilutions <- function(
  max.conc.std = max.conc.std, n.dilutions = n.dilutions, dilution.fact = dilution.fact) {
  dilutions <- sapply(
    seq(0, (n.dilutions - 1)),
    function(x) max.conc.std / (dilution.fact^x)
  )
  return(dilutions)
}

# Get background-corrected standard
# TODO: add documentation
get.standard <- function(data = data,
                         row.idx = row.idx.std,
                         col.idx = col.idx.std,
                         n.dilutions = n.dilutions,
                         n.repl = n.repl,
                         dilution.fact = dilution.fact,
                         max.conc.std = max.conc.std, 
                         dilutions = dilutions) {
  row.end <- row.idx + (n.dilutions * 2) - 1
  col.end <- col.idx + (n.repl - 1)
  standard <- data %>% .[row.idx:row.end, col.idx:col.end]
  corr.standard <- purrr::map_df(
    seq(1, row.end, by = 2),
    ~ correct.sample(data = standard, ., col.idx = 1, n.cols = n.repl)
  ) %>%
    na.omit() %>%
    setNames(seq(n.repl)) %>%
    dplyr::mutate(dilution = dilutions) %>%
    tidyr::gather(key = replicate, value = absorbance, -dilution)

  return(corr.standard)
}

# TODO: add documentation
get.donors <- function(row.idx = row.idx.donor.1,
                       col.idx = col.idx.donor.1,
                       n.cols = length(tpoints),
                       n.repl = n.repl.donors) {
  row.end <- row.idx + 1 + 2 * (n.repl.donors - 1) # here, replicates are performed row-wise
  col.end <- col.idx + (n.cols - 1)
  sample <- data %>% .[row.idx:row.end, col.idx:col.end]
  corr.donors <- correct.sample(
    data = sample, row.idx = 1, col.idx = 1, n.cols = length(tpoints)
  )
  return(corr.donors)
}
```

Now we can create a data frame with the corrected data of the standard:
```{r, get-standard, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE}
dilutions <- comp.dilutions(
  max.conc.std = max.conc.std,
  n.dilutions = n.dilutions,
  dilution.fact = dilution.fact
)

standard <- get.standard(
  data = data,
  row.idx = row.idx.std,
  col.idx = col.idx.std,
  n.dilutions = n.dilutions,
  n.repl = n.repl.std,
  dilution.fact = dilution.fact,
  max.conc.std = max.conc.std, 
  dilutions = dilutions
)
# Inspect result
standard
```

Here we select the wells with the blanks:
```{r, get-blanks, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE}
# get blanks
blanks <- correct.sample(data, row.idx.blank, col.idx.blank, n.cols = 2) %>%
  setNames(seq(1, dim(.)[2])) %>%
  tidyr::gather(key = replicate, value = blank)
# Inpsect blanks
blanks
```

And finally we create a data frame with the corrected data of our biological samples. In our case, we have a time series with five data points for four donors:
```{r, select-samples, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE}
donors <- purrr::map_df(
  seq(row.idx.donor.1, (row.idx.donor.1 + 2 * (n.donors - 1)), by = 2),
  ~ get.donors(., col.idx = col.idx.donor.1, n.cols = length(tpoints), n.repl = n.repl.donors)
) %>%
  setNames(tpoints) %>%
  dplyr::mutate(donor = seq(1, n.donors)) %>%
  dplyr::mutate(donor = factor(donor, unique(donor))) %>%
  tidyr::gather(key = "time", value = "absorbance", -donor) %>%
  dplyr::mutate(time = as.integer(time))

# Inspect data
donors
```

Before we proceed with the analysis, let's plot the data and see whether it looks reasonable. 