# Plot data {#plot-data}
For convience, we define a plot theme that suits our taste:

## Define plot theme
```{r, plot-theme, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, cache = TRUE}
min.absorb <- min(blanks$blank)
max.absorb <- max(standard$absorbance)
font.size <- 12
# Define theme for plotting
#' @param title.hjust horizontal alignment of plot title
#' @param legend_pos legend position
#' @return theme for ggplot
theme_plot <- function(
  title.hjust = 0, legend.pos = "right", legend.dir = "vertical") {
  theme(
    axis.text = element_text(
      size = font.size,
      face = "bold"
    ),
    axis.title.x = element_text(
      size = font.size,
      hjust = 1
    ),
    axis.title.y = element_text(
      size = font.size,
      hjust = 0.9
    ),
    plot.title = element_text(
      size = font.size + 2,
      face = "bold",
      hjust = title.hjust
    ),
    plot.margin = rep(grid::unit(1, "cm"), 4),
    strip.text.x = element_text(size = font.size),
    strip.text.y = element_text(size = font.size),
    strip.background = element_rect(fill = "white"),
    axis.line = element_blank(),
    legend.position = legend.pos,
    legend.direction = legend.dir,
    legend.text = element_text(size = font.size),
    legend.title = element_text(size = font.size)
  ) +
    background_grid(
      major = "yx",
      minor = "",
      colour.major = "grey90",
      size.major = 0.2
    )
}
```

## Plot standard
Now let's plot the standard:
```{r, plot-std, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, cache = TRUE, fig.height = 4}
plot.std <- function(semilog = FALSE) {
  p <- ggplot(data = standard) +
    scale_color_brewer(
      palette = "Set1",
      labels = paste("Replicate", seq(n.repl.std)),
      guide = guide_legend(title = "")
    ) +
    scale_y_continuous(limits = c(min.absorb, max.absorb + 0.1 * max.absorb)
    ) +
    geom_hline(
      yintercept = mean(blanks$blank),
      lty = 2,
      color = "grey60",
      lwd = 0.2
    ) +
    theme_plot()

  if (semilog == TRUE) {
    p <- p + geom_point(aes(log2(dilution), absorbance, color = replicate), alpha = 0.6) +
      labs(
        x = bquote(log[2] * " HI titer (w/o dilution factor 1:5000)"),
        y = "Absorbance (a. u.)",
        title = "IgG in standard serum"
      ) +
      scale_x_continuous(breaks = log2(dilutions)) +
      annotate("text",
        label = "Blank",
        x = max(log2(standard$dilution)) - 0.05 * max(log2(standard$dilution)),
        y = max(blanks$blank) + 20 * max(blanks$blank),
        size = 3.5,
        color = "grey50"
      )
  }
  else {
    p <- p + geom_point(
      aes(dilution, absorbance, color = replicate),
      alpha = 0.6
    ) +
      labs(
        x = "HI titer (w/o dilution factor 1:5000)",
        y = "Absorbance (a.u.)",
        title = "IgG in standard serum"
      ) +
      scale_x_continuous(breaks = dilutions) +
      annotate("text",
        label = "Blank",
        x = max(standard$dilution) - 0.05 * log2(standard$dilution),
        y = max(blanks$blank) + 0.2 * max(blanks$blank),
        size = 3.5,
        color = "grey50"
      )
  }
  p
}
plot.std()
```

We see that the concentration of the dilution series is on a logarithmic scale. Let's replot the results with $\log_2$-transformed concentrations ($\log_2$ for a dilution factor of 2):
```{r, plot-std-log, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, cache = TRUE, fig.height = 4}
plot.std(semilog = TRUE)
```

## Plot biological samples
Now let's have a look at our actual data. We measured IgG in serum of four donors on five different time points. We first define a plot for our biological samples and then plot it together with the standard curve to have a better overview.
```{r, plot-samples, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, chache = TRUE}
plot.donors <- ggplot(
    data = donors,
    aes(time, absorbance, color = donor, id = donor)
  ) +
    geom_point(alpha = 0.7) +
    geom_smooth(alpha = 0.5, lwd = 0.5, se = FALSE) +
    labs(
      x = "Days after vaccination", y = "Absorbance (a. u.)",
      title = "IgG in serum of healthy volunteers"
    ) +
    scale_color_viridis(
      begin = 0.2, end = 0.8, discrete = TRUE,
      labels = paste("Donor", seq(n.donors)), guide = guide_legend(title = "")
    ) +
    scale_y_continuous(limits = c(min.absorb, max.absorb + 0.1 * max.absorb)) +
    scale_x_continuous(breaks = tpoints) +
    geom_hline(
      yintercept = mean(blanks$blank),
      lty = 2, color = "grey60", lwd = 0.2
    ) +
    annotate("text",
      label = "Blank",
      x = max(donors$time) - 0.1 * max(donors$time),
      y = max(blanks$blank) + 20 * max(blanks$blank),
      size = 3.5, color = "grey50"
    ) +
    background_grid(
      major = "yx", minor = "", colour.major = "grey90", size.major = 0.2
    ) +
    theme_plot()
```

## Combine plots
```{r, plot-grid, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, cache = TRUE, fig.height = 8}
plot_grid(plot.std(semilog = TRUE), plot.donors, ncol = 1)
```