# Plot data {#plot-data}
For convenience, we first define a plot theme:

## Define plot theme
```{r, plot-theme, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, cache = TRUE}
min.absorb <- min(blanks$blank)
max.absorb <- max(standard$absorbance)
font.size <- 11
# Define theme for plotting
#' @param title.hjust horizontal alignment of plot title
#' @param legend_pos legend position
#' @return theme for ggplot
theme_plot <- function(
  title.hjust = 0, legend.pos = "right", legend.dir = "vertical") {
  theme(
    axis.text = element_text(
      size = font.size,
      face = "bold"
    ),
    axis.title.x = element_text(
      size = font.size,
      hjust = 1
    ),
    axis.title.y = element_text(
      size = font.size,
      hjust = 0.9
    ),
    plot.title = element_text(
      size = font.size + 2,
      face = "bold",
      hjust = title.hjust
    ),
    plot.margin = rep(grid::unit(1, "cm"), 4),
    axis.line = element_blank(),
    legend.position = legend.pos,
    legend.direction = legend.dir,
    legend.text = element_text(size = font.size),
    legend.title = element_text(size = font.size)
  ) +
    background_grid(
      major = "yx",
      minor = "",
      colour.major = "grey90",
      size.major = 0.2
    )
}
```

## Plot standard
Now let's plot the standard:
```{r, plot-std, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, cache = TRUE, fig.height = 4}
plot.std <- function(semilog = FALSE) {
  p <- ggplot(data = standard) +
    scale_color_brewer(
      palette = "Set1",
      labels = paste("Replicate", seq(n.repl.std)),
      guide = guide_legend(title = "")
    ) +
    scale_y_continuous(
      limits = c(min.absorb, max.absorb + 0.1 * max.absorb)
    ) +
    geom_hline(
      yintercept = mean(blanks$blank),
      lty = 2,
      color = "grey60",
      lwd = 0.2
    ) +
    theme_plot()

  if (semilog == TRUE) {
    p <- p + geom_point(aes(concentration, absorbance, color = replicate), alpha = 0.6) +
      labs(
        x = "Relative concentration",
        y = "Absorbance (a. u.)",
        title = "IgG in standard serum"
      ) +
      scale_x_continuous(breaks = concentrations, labels = scales::percent, trans = "log2") +
      annotate("text",
        label = "Blank",
        x = max(standard$concentration) - 0.05 * max(standard$concentration),
        y = max(blanks$blank) + 20 * max(blanks$blank),
        size = 3.5,
        color = "grey50"
      )
  }
  else {
    p <- p + geom_point(
      aes(concentration, absorbance, color = replicate),
      alpha = 0.6
    ) +
      labs(
        x = "Relative concentration",
        y = "Absorbance (a.u.)",
        title = "IgG in standard serum"
      ) +
      scale_x_continuous(breaks = round(concentrations,1),
                         labels = scales::percent) +
      annotate("text",
        label = "Blank",
        x = max(standard$concentration) - 0.05 * max(standard$concentration),
        y = max(blanks$blank) + 0.2 * max(blanks$blank),
        size = 3.5,
        color = "grey50"
      )
  }
  p
}
plot.std()
```

We see that the concentration of the dilution series is on a logarithmic scale. Let's replot the results with $\log_2$-transformed concentrations ($\log_2$ for a dilution factor of 2):
```{r, plot-std-log, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, cache = TRUE, fig.height = 4}
plot.std(semilog = TRUE)
```

## Plot biological samples
Now let's have a look at our actual data. We measured IgG in serum of four donors on five different time points. We first define a plot for our biological samples and then plot it together with the standard curve to have a better overview.
```{r, plot-samples, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, chache = TRUE}
# select HV
ID.selected <- c("BD1989", "MS1984", "SM1988", "HJ1992")

donors.av <- donors.av %>% dplyr::mutate(ID = c(rep("BD1989", 5), rep("MS1984", 5), rep( "SM1988", 5), rep("HJ1992", 5)))
donors.av
plot.donors <- ggplot() +
    geom_point(data = donors.av, 
      aes(time, absorbance.av, color = donor), alpha = 0.7) +
    geom_line(data = donors.av, 
      aes(time, absorbance.av, group = donor, color = donor), alpha = 0.6, lwd = 0.5) +
    labs(
      x = "Days after vaccination", y = "Absorbance (a. u.)",
      title = "IgG in serum of healthy volunteers"
    ) +
    scale_color_viridis(
      begin = 0.2, end = 0.8, discrete = TRUE,
      labels = paste("Donor", seq(n.donors)), guide = guide_legend(title = "")
    ) +
    scale_y_continuous(limits = c(min.absorb, max.absorb + 0.1 * max.absorb)) +
    scale_x_continuous(breaks = tpoints) +
    geom_hline(
      yintercept = mean(blanks$blank),
      lty = 2, color = "grey60", lwd = 0.2
    ) +
    annotate("text",
      label = "Blank",
      x = max(donors$time) - 0.1 * max(donors$time),
      y = max(blanks$blank) + 20 * max(blanks$blank),
      size = 3.5, color = "grey50"
    ) +
    background_grid(
      major = "yx", minor = "", colour.major = "grey90", size.major = 0.2
    ) +
    theme_plot() + ylim(0,2)

# select HV
ID.selected <- c("BD1989", "MS1984", "SM1988", "HJ1992")
titer.s <- titer %>%
  dplyr::mutate(ID = as.character(ID)) %>%
  dplyr::filter(virus == "California") %>%
  dplyr::filter(ID %in% ID.selected) 

p.titer <- ggplot(data = titer.s) +
  geom_point(data=titer.s, aes(time, log2titer, color = ID), alpha = 0.7) +
  geom_line(
    data = titer.s,
    aes(time, log2titer, group = ID, color = ID), alpha = 0.6, lwd = 0.5
  ) +
  labs(
    x = "Days after vaccination", y = "log2 HI titer",
    title = "Corresponding HI titer"
  ) +
  scale_color_viridis(
    begin = 0.2, end = 0.8, discrete = TRUE, guide = FALSE
  ) +
  scale_y_continuous(limits = c(0, max(titer.s$log2titer) + 0.1 * max(titer.s$log2titer))) +
  scale_x_continuous(breaks = tpoints) +
  background_grid(
    major = "yx", minor = "", colour.major = "grey90", size.major = 0.2
  ) +
  theme_plot() +
  panel_border()
```

## Combine plots
```{r, plot-grid, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, cache = TRUE, fig.height = 8}
plot_grid(plot.std(semilog = TRUE), plot.donors, ncol = 1)
```