# Estimate calibration curve {#calibration}
## The four-parameter logistic curve
The relationship between concentration and absorbance can be described by the so called _four-parameter logistic equation (4-PL)_. We will use the following equation:

$$f(c) =  \mathrm{A_{min}} + \frac{\mathrm{A_{max}} - \mathrm{A_{min}}}{1+\exp(\mathrm{n} \cdot (\log_2(c)-\log_2(\mathrm{EC}_{50})))}$$,

where $\mathrm{A}$ stands for absorbance and $c$ for concentration. The parameters that we are going to estimate are $\mathrm{A_{max}}$, which is the maximal absorbance that can be obtained (for infinite concentration $c \rightarrow \infty$), $\mathrm{A_{min}}$, which is the minimum absorbance that can be obtained (for $c \rightarrow 0$), and parameters $\mathrm{EC}_{50}$ and $n$. The value of $\log \mathrm{EC}_{50}$ is the inflection point of the curve. This means, $\log \mathrm{EC}_{50}$ is the concentration for which we obtain 50% of $\mathrm{A_{max}}$. Parameter $n$ is related to the steepness of the curve at the inflection point. 

For convenience, we use $\log_2$ in the equation (and not $log_{10}$ as often used in literature), because then the estimated $\log_2 EC_{50}$ is easier to interpret, since we performed a 1:2 dilution series to generate the concentrations of our standard. Note that $\log_2(x) = \frac{log_{10}(x)}{log_2(10)} = 0.30103 \cdot \log_{10}(x)$. Changing the base simply scales all concentration by a constant value (linear transformation).

## Regression of the standard
We fit the data obtained for the standard to the four-parameter logistic curve. This is a non-linear least squares problem and we use the  Levenberg-Marquadt algorithm to solve it. 
The Levenberg-Marquardt algorithm is more robust than the Gauss–Newton algorithm (used by default by the `nls()` function of the `nls` package), which means it converges to a minimum even if the initial starting values are far from the final minimum. 
```{r, test-fit, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, cache = TRUE}
# Fit 4-parameter logistic dose-response curve
# Approach 1): Use nlsLM pacakge for Levenberg-Marquardt optimization
levm.fit <- nlsLM(
  absorbance ~
  Amin + ((Amax - Amin)) / (1 + exp(n * (log2(dilution) - log2EC50))),
  data = standard,
  start = list(Amin = 0, Amax = 2, log2EC50 = 8, n = -1)
)
# Show summary of the fit
summary(levm.fit)
```
Apparently, we can not estimate $\mathrm{A_{min}}$ from the data. But we can fix it to a certain value: we know that $\mathrm{A_{min}}$ can not be negative and we can use the values obtained from the blank wells as an estimate for $\mathrm{A_{min}}$.

```{r, lm-fit, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, fig.height=4}
# Set Amin to blank values
Amin <- median(blanks$blank)
# Re-fit with only three free parameters
levm.fit <- nlsLM(
  absorbance ~
  Amin + ((Amax - Amin)) / (1 + exp(n * (log2(dilution) - log2EC50))),
  data = standard,
  start = list(Amax = 2, log2EC50 = 8, n = -1)
)
# Check new fit results
summary(levm.fit)
# Show 95% confidence interval
confint(levm.fit)
## Add calibration curve to plot
# Save estimated parameters
Amax.est <- coef(levm.fit)[[1]]
n.est <- coef(levm.fit)[[2]]
log2EC50.est <- coef(levm.fit)[[3]]
# Full range of considered concentrations
dilution.new <- seq(2^4, 2^11)
# Predict new absorbance values from calibration curve
calibration.curve <- tibble(
  dilution.new = dilution.new,
  absorbance.est = predict(levm.fit, tibble(dilution = dilution.new))
)

## Compute prediction intervals
# TODO: double-check, ignore this part for now
fgh <- deriv(
  absorbance ~ Amin + ((Amax - Amin)) / (1 + exp(n * (log2(dilution) - log2EC50))),
  c("Amax", "n", "log2EC50"), function(Amax, n, log2EC50, dilution) {}
)
f.new <- fgh(Amax.est, n.est, log2EC50.est, dilution.new)
g.new <- attr(f.new, "gradient")
cov.fit <- vcov(levm.fit)
gs <- rowSums((g.new %*% cov.fit) * g.new)
alpha <- 0.05
sigma.est <- summary(levm.fit)$sigma
delta.y <- sqrt(gs + sigma.est^2) * qt(1 - alpha / 2, 15)
calibration.curve <- calibration.curve %>%
  dplyr::mutate(pred.lwr = absorbance.est - delta.y) %>%
  dplyr::mutate(pred.upr = absorbance.est + delta.y)

## Plot calibration curve
plot.ccurve <- plot.std(semilog = TRUE) +
  geom_line(data = calibration.curve,
    aes(log2(dilution.new), absorbance.est), alpha = 0.6) + 
  geom_ribbon(data = calibration.curve,
    aes(x = log2(dilution.new), ymin = pred.lwr, ymax = pred.upr),
    alpha = 0.2, fill = "grey") +
  scale_y_continuous(limits = c(-0.1, 2))
plot.ccurve

# Plot predicted absorbance vs residuals
plot(levm.fit)
```

Alternatively, one can use the `drc` package to fit a 4-PL curve using. However, the `drm` function performs a $\log_10(x)$ transformation by itself, so one get's results for $\log_10$-transformed concentrations.
```{r, drc-fit, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, cache = TRUE, fig.height=4}
# Fit 4-parameter logistic dose-response curve
# Approach 2): Use drc package
# NOTE: the drm function performs a log10(x) transformation by itself.
# See also ?LL2.4 for more information
drc.fit <- drm(absorbance ~ dilution,
  data = standard,
  fct = LL2.4(names = c("n", "Amin", "Amax", "EC50"))
)
summary(drc.fit)
plot(drc.fit)
ggplot(as.tibble(drc.fit$predres)) +
  geom_point(aes(`Predicted values`, Residuals),
    pch = 1, size = 3, color = "cornflowerblue") +
  theme_plot() + 
  panel_border() + 
  geom_hline(yintercept = 0, lwd = 0.2)
```


  
  
