---
title: "Analyse ELISA data in _R_"
author: "JL"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: janalin/ELISA-analysis
description: ""
---

# Introduction {#intro}
This document shows how to analyse ELISA data in _R_. The code was written to analyse ELISA experiments performed in our lab: the script reads the excel file provided by a _Synergy H1_ 96-well plate reader from BioTek, performs a background-correction, generates the calibration curve and computes the sample concentrations with respect to this calibration curve.

## About the experiment in brief
We measured the IgG amount in blood serum of four donors using an _indirect ELISA_. The detection antibody was conjugated to horseradish peroxidase (HRP). As a substrate solution, we used TMB from Thermo Fisher Scientific. We measured absorbance at 450 nm -- to measure the substrate specific absorbance -- and at 620 nm -- to measer the (unspecific) background absorbance. The absorbance of both wavelength is provided by the plate reader in a single excel sheet. If we say background correction we mean that we substract the values obtained at 620 nm from the values obtained at 450 nm.

## Load libraries and import excel fiele
We first load some _R_ libraries that we want to use.
```{r, load-libs, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE}
# load packages
if (!requireNamespace("pacman", quietly = TRUE)){
  install.packages("pacman")}
pacman::p_load(bookdown, # to create this document
               tidyverse, # for data wrangling
               ggplot2, cowplot, # for nice plots
               gdata, # to read excel files
               here, # to find the absolute path of the project
               purrr, # for functional programming
               viridis, # for color palettes
               drc, minpack.lm) # for 4-param logistic regression
```

Now we read the excel file:
```{r, import-data, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE}
# Get the project path
path <- here::here()
# Set the name of the excel file
file <- "/../data/180821-IgG_HV_Test.xlsx"
# Import excel data
data <- gdata::read.xls(paste0(path, file), sheet = 1, header = FALSE) %>% as.tibble
# Let's have a look at the absorbance data.
head(data[complete.cases(data),])
```

The odd rows show the absorbance at 450 nm and the even rows the absorbance at 620 nm. As expected, the background absorbance at 620 nm is very low. 

And these are our plate reader specifications
```{r, plate-reader, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE}
# Inspect plate reader specifications:
data[,1:2]
```

## Set cell indices and set experimental parameteres
We need to specify some experiment-specific parameters manually such as the position of the wells with the standard samples and blanks, the number of replicates, etc. 
```{r, select-cells, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, cache = TRUE}
# Use View(data) to double-check row and column indices

## BLANKS
# Location of the 450 nm measurment of our blanks
row.idx.blank <- 36
col.idx.blank <- 13

## STANDARD SAMPLES
# Location of the 450 nm measurment of the first well 
# of our standard (well 1A in our case)
row.idx.std <- 22
col.idx.std <- 3
# Number of replicates of the standard
n.repl.std <- 2
# Maximal concentration of the standard 
max.conc.std <- 2^11
# Number of dilutions performed to obtain the calibration curve
n.dilutions <- 8
# Dilution factor
dilution.fact = 2

## BIOLOGICAL SAMPLES
# Location of first well with the sample of the first donor (here well 3A)
row.idx.donor.1 <- 22
col.idx.donor.1 <- 5
n.donors <- 4
tpoints <- c(0, 7, 30, 60, 180)
n.repl.donors <- 1
```

Now we can start analysing the data.