---
title: "Analyse ELISA data in _R_"
author: "JL"
date: "2018-11-27"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: janalin/ELISA-analysis
description: ""
---

# Introduction {#intro}
This document shows how to analyse ELISA data in _R_. The code was written to analyse ELISA experiments performed in our lab: the script reads the excel file provided by a _Synergy H1_ 96-well plate reader from BioTek, performs a background-correction, generates the calibration curve and computes the sample concentrations with respect to this calibration curve.

Note from 04-Apr-2024: This document was developed in 2018 and may contain outdated and partially experimental code -- use on your own responsibility. You should understand the calculations behind the analysis before applying them to your data.

## About the experiment in brief
We measured the IgG amount in blood serum of four donors using an _indirect ELISA_. The detection antibody was conjugated to horseradish peroxidase (HRP). As a substrate solution, we used TMB from Thermo Fisher Scientific. We measured absorbance at 450 nm -- to measure the substrate specific absorbance -- and at 620 nm -- to measure the (unspecific) background absorbance. The absorbance of both wavelength is provided by the plate reader in a single excel sheet. If we say background correction we mean that we substract the values obtained at 620 nm from the values obtained at 450 nm.

## Load libraries and import excel file
We first load some _R_ libraries that we want to use.
```{r, load-libs, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE}
# load packages
if (!requireNamespace("pacman", quietly = TRUE)){
  install.packages("pacman")}
pacman::p_load(bookdown, # to create this document
               tidyverse, # for data wrangling
               ggplot2, cowplot, # for nice plots
               scales, # for axis scales with '%'
               readxl, # to read excel files
               here, # to find the absolute path of the project
               purrr, # for functional programming
               viridis, # for color palettes
               drc, minpack.lm) # for 4-param logistic regression
```

Now we read the excel file:
```{r, import-data, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE}
# Set the name of the excel file
# Note: this is example data used for illustrative purposes only.
xlsx.file <- file.path(here::here(), "example-data.xlsx")
# Import excel data
data <- readxl::read_xlsx(xlsx.file, sheet = 1, col_names = FALSE) %>% as.tibble
# Let's have a look at the absorbance data.
# View(data)
```

The odd rows show the absorbance at 450 nm and the even rows the absorbance at 620 nm. As expected, the background absorbance at 620 nm is very low. 

## Set cell indices and set experimental parameteres
We need to specify some experiment-specific parameters manually such as the position of the wells with the standard samples and blanks, the number of replicates, etc. 
```{r, select-cells, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, cache = TRUE}
## BLANKS
# Location of the 450 nm measurement of our blanks in the `data` table.
# Use View(data) to double-check row and column indices:
row.idx.blank <- 40
col.idx.blank <- 13

## STANDARD SAMPLES
# Location of the 450 nm measurement of the first well 
# of our standard (well 1A in our case -> row 26 and column 3 in `data` table)
row.idx.std <- 26
col.idx.std <- 3
# Number of replicates of the standard
n.repl.std <- 2
# Maximal concentration of the standard 
max.conc.std <- 1 # i.e. 100 %, we use a relative scale
# Number of dilutions performed to obtain the calibration curve
n.dilutions <- 8
# Dilution factor
dilution.fact = 2

## BIOLOGICAL SAMPLES
# Location of first well with the sample of the first donor (here well 3A)
row.idx.donor.1 <- 26
col.idx.donor.1 <- 5
n.donors <- 4 # Four donors, pipetted in horizontal order
tpoints <- c(0, 7, 30, 60, 180) # Five time points, pipetted in vertical order
n.repl.donors <- 2 # Two replicate measurements per donor and time point
```

Now we can start analysing the data.